---
title: "Final Report"
author: "Jacinda Chen"
date: "March 1, 2020"
bibliography: references.bib
csl: ecology.csl
output: word_document
---

```{r setup, include=FALSE}
# Set up
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Delete this placeholder text, and replace it with your introduction text.

# Methods

# Add text here.

## Site Locations and Descriptions

Add text here.

## Field Sampling Design

# Add text here.

## Data Analysis and Statistics

# Add text here.

# Results

## Subsections are ok in the results section too

# Add a number of code chunks in the Results section. These should read in, subset and plot the data as needed (no need to save any figures to pdf, since they will be put into the rendered document when you click 'knit'), and, for any hypotheses that you want to test, an appropriate statistical test. 

```{r load-packages}
# Load libraries
library("dplyr")
library("tidyr")
library("ggplot2")
library("readr")
library("rebird")
library("citr")
```

# Discussion

# Sources Cited

